### 2012-07-30 Doc/VRIPHYS2012/vriphys2012.tex
small changes in section 3
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@439 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

Thus, we can locally compute some spacial derivatives, along $x$ and $y$ on the interpolation function vector $T$ (which has three components $(T^x, T^y, T^z)$ .
The 2D tensor $F$ of the in-plane deformations can be computed as a $2\times2$ matrix where the line $k$ and the column $l$ alternatively represent $x$ and $y$\footnote{It can be noted $k=\{x,y\}$ and $l=\{x,y\}$}:
\begin{equation}
F_{kl} = \frac{\partial T^k}{\partial l}  =  \sum_{i=1}^3 \frac{\partial T^k}{\partial \xi_i} \frac{\partial \xi_i}{\partial l}  
\end{equation}

The term $\frac{\partial T^k}{\partial \xi_i}$ can be derived from the expression of $T(\xi_1,\xi_2, \xi_3)$ in equation (\ref{global-transform}) and 
$\frac{\partial \xi_i}{\partial l}$ from equation (\ref{shape-vector}). 
The in-plane displacement gradient can be obtained with tensor $F$:
\begin{equation}
\frac{\partial u^k}{\partial l} = F_{kl} - I_{kl} 
\end{equation}
where $I$ is the identity matrix ($I_{kl}  = \{1$ if $l=k$ ;  $0$ if $l \neq k \} $). 
The in-plane displacement gradient can be linked to the degrees of freedom of the shell using the expression of the displacement of a point $\delta T$ (see equation  (\ref{bez-delta})) on the shell surface:
$$
\frac{\partial u^k}{\partial l} = \frac{\partial (\delta T^k)}{\partial l}
$$

@@

    \item Elastic membrane defining deformations in plane of the
    element. It encompasses deformations like stretching and shearing.
    It corresponds to the in-plane displacements $u_x, u_y$. 
    At the nodes' level, it corresponds to in-plane displacements $\delta\mathbf{x}_x, \delta\mathbf{x}_y$ and rotation $\theta_z$.
(Figure \ref{fig-membrane}).
    \item Bending plate defining deformation due to out of plane bending displacement $u_z$. It is
    described by 3 degrees of freedom: rotation around two axes $\theta_x,
    \theta_z$ and out of plane displacement $\delta\mathbf{x}_z$ (Figure \ref{fig-plate}).

@@

To compute the values in matrices $\left[ \mathbf{J}_m \right]$ and $\left[ \mathbf{J}_b \right]$,  we can reuse equation (\ref{bez-delta}) and apply the spacial derivatives.
To illustrate, let's take one value of the matrix $\left[ \mathbf{J}_m \right]$  (first line, fourth column), we first take the kinematic between $\delta T_x$ and the concerned degree of freedom $ \delta \mathbf{x}_{2x}$:
$$
   \frac{ \delta T_x}{ \delta \mathbf{x}_{2x}} = (n_2 + n_6 + n_7 + \frac{1}{3} n_{10})
$$
Then, we apply the spacial derivatives to compute the variation of $\epsilon_{xx}$ when 
$$
\delta \epsilon_{xx} += \underbrace{\sum_{i=1}^{3} (\frac{\partial n_2}{\xi_i} + \frac{\partial n_6}{\xi_i} + \frac{\partial n_9}{\xi_i} + \frac{1}{3}\frac{\partial n_{10}}{\xi_i})\frac{\partial \xi_i}{\partial x}}_{\left[ \mathbf{J}_m \right](1,4)}  {\delta \mathbf{x}_{2x}}
$$

@@

Because the deflection field for the element is non-linear \eqref{eq-delta}
in position the integrals \eqref{eq-Km} and \eqref{eq-Kb} have to be computed using
numerical integration. In our implementation we employ 6-point Gaussian
quadrature for integration over triangle area.
% TODO: add ref
% TODO: add abscissas/weights?

>>>>

Thus, we can locally compute some spacial derivatives, along $x$ and $y$ on the interpolation function vector $T$ (which has three components $(T^x, T^y, T^z)$) .
The 2D tensor $F$ of the in-plane deformations can be computed as a $2\times2$ matrix where the line $k$ and the column $l$ alternatively represent $x$ and $y$\footnote{It can be noted $k=\{x,y\}$ and $l=\{x,y\}$}:
\begin{equation}
F_{kl} = \frac{\partial T^k}{\partial l}  =  \sum_{i=1}^3 \frac{\partial T^k}{\partial \xi_i} \frac{\partial \xi_i}{\partial l}  
\end{equation}

The term $\frac{\partial T^k}{\partial \xi_i}$ can be derived from the expression of $T(\xi_1,\xi_2, \xi_3)$ in equation (\ref{global-transform}) and 
$\frac{\partial \xi_i}{\partial l}$ from equation (\ref{shape-vector}). 
The in-plane displacement gradient can be obtained with tensor $F$:
\begin{equation}
\frac{\partial u^k}{\partial l} = F_{kl} - I_{kl} 
\end{equation}
where $I$ is the identity matrix ($I_{kl}  = \{1$ if $l=k$ ;  $0$ if $l \neq k \} $). 
%
%
% Remark: the sentence can be suppr..
\CD{The following  sentencd I let you chose: keep or not...}The in-plane displacement gradient can be linked to the degrees of freedom of the shell using the expression of the displacement of a point $\delta T$ (see equation  (\ref{bez-delta})) on the shell surface:
$$
\frac{\partial u^k}{\partial l} = \frac{\partial (\delta T^k)}{\partial l}
$$ 
% end of Remark

@@

    \item Elastic membrane defining deformations in plane of the
    element. It encompasses deformations like stretching and shearing.
    It corresponds to the in-plane displacements $u_x, u_y$. 
    At the nodes' level, it corresponds to in-plane displacements $\delta\mathbf{x}_x, \delta\mathbf{x}_y$ and rotation $\theta_z$.
(Figure \ref{fig-membrane}).
    \item Bending plate defining deformation due to out of plane bending displacement $u_z$. It is
    influenced by 3 degrees of freedom: rotation around two axes $\theta_x,
    \theta_z$ and out of plane displacement $\delta\mathbf{x}_z$ (Figure \ref{fig-plate}).

@@

To compute the values in matrices $\left[ \mathbf{J}_m \right]$ and $\left[ \mathbf{J}_b \right]$,  we can reuse equation (\ref{bez-delta}) and apply the spacial derivatives.
To illustrate, let's take one value of the matrix $\left[ \mathbf{J}_m \right]$  (first line, fourth column), we first take the kinematic between $\delta T_x$ and the concerned degree of freedom $ \delta \mathbf{x}_{2x}$ from equation (\ref{bez-delta}):
$$
   \frac{ \delta T_x}{ \delta \mathbf{x}_{2x}} = (n_2 + n_6 + n_7 + \frac{1}{3} n_{10})
$$
Then, we apply the spacial derivatives to compute the variation of $\epsilon_{xx}$ due to a variation of $\mathbf{x}_{2x}$.
$$
\delta \epsilon_{xx} += 
\underbrace{\left( \sum_{i=1}^{3} (\frac{\partial n_2}{\xi_i} + \frac{\partial n_6}{\xi_i} + \frac{\partial n_7}{\xi_i} + \frac{1}{3}\frac{\partial n_{10}}{\xi_i})\frac{\partial \xi_i}{\partial x} \right) }_{\left[ \mathbf{J}_m \right](1,4)}  
{\delta \mathbf{x}_{2x}}
$$

@@

Because the deflection field for the element is non-linear \eqref{eq-delta}
in position the integrals \eqref{eq-Km} and \eqref{eq-Kb} have to be computed using
numerical integration. In our implementation we employ 6-point Gaussian
quadrature for integration over triangle area \CD{TODO: figure + gives the formula of the quadrature...}
% TODO: add ref
% TODO: add abscissas/weights?
